---
title: "STAT 5014 Homework 5"
author: "Samantha Sunshine"
date: "October 3, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(ggplot2)
```

# Problem 3

If the figure is of data represented in a table, it should contain tidy data with proper labels on the rows and columns. If the figure is a graph, it should have proper axis labels, color and/or shading corresponding to different parts of the graph, a visually appealing layout, and an easy way to interpret what the plot is trying to show for people who might not be familiar with the subject.

# Problem 4

```{r, echo = FALSE}
#computes the proportion of successes in a vector consisting of 0s and 1s
prop_success <- function(x){
  sum(x)/length(x)
}

#matrix simulating 10 flips of a coin
set.seed(12345)
P4b_data <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10, ncol = 10)

#apply computes the proportion of successes in each column
apply(P4b_data, 2, prop_success)

#apply computes the proportion of successes in each row
apply(P4b_data, 1, prop_success)

#given an input of a probability, rbinom gives a vector with outcomes of 10 coin flips
prob <- function(x){
  rbinom(10, 1, prob = x)
}

#vector of probabilities between .3 and .4
prob_vec <- seq(0.3, 0.39, 0.01)

#sapply gives a matrix of 0s and 1s with the appropriate probabilities
prob_mat <- sapply(prob_vec, prob)

#computes the marginal successes for each column of the matrix
prop_success <- function(x){
  sum(x)/length(x)
}

apply(prob_mat, 2, prop_success)
```

The proportion of successes in each column is the same, 0.6. The proportion of successes in each row is either 0 or 1. 
The last row is the vector of marginal success of each column of our matrix.

# Problem 5

```{r, echo = FALSE}
url <- "http://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat"
starch_raw <- read.table(url, header = T, stringsAsFactors = F)

#plot of starch hickness vs. strength
plot(starch_raw$thickness, starch_raw$strength, xlab = "Thickness", ylab = "Strength", main = "Starch Thickness vs, Strength")

#separating each type of starch
starch_ca <- starch_raw[1:13, ]
starch_co <- starch_raw[14:32, ]
starch_po <- starch_raw[33:49, ]

#means of strength each type of starch
strength_ca_mean <- mean(starch_ca$strength)
strength_co_mean <- mean(starch_co$strength)
strength_po_mean <- mean(starch_po$strength)
strength_means <- data.frame(c("CA", "CO", "PO"), c(strength_ca_mean, strength_co_mean, strength_po_mean))
colnames(strength_means) <- c("starch", "strength")

#means thickness of each type of starch
thickness_ca_mean <- mean(starch_ca$thickness)
thickness_co_mean <- mean(starch_co$thickness)
thickness_po_mean <- mean(starch_po$thickness)
thickness_means <- data.frame(c("CA", "CO", "PO"), c(thickness_ca_mean, thickness_co_mean, thickness_po_mean))
colnames(thickness_means) <- c("starch", "thickness")

#bar plots of the means in strength and thickness of each starch
par(mfrow = c(2, 2))
ggplot(strength_means, aes(x = starch, y = strength)) + geom_bar(stat = "identity") + ggtitle("Mean Strength of Different Starches")
ggplot(thickness_means, aes(x = starch, y = thickness)) + geom_bar(stat = "identity") + ggtitle("Mean Thickness of Different Starches")
```

# Problem 6

```{r, echo = FALSE}
#we are grabbing a SQL set from here
#http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip

# download the files, looks like it is a .zip
library(downloader)
download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip", dest = "us_cities_states.zip")
unzip("us_cities_states.zip", exdir = "C:/Users/ssuns/Documents/2017 Fall/Statistical Programming/STAT_5014_Homework")

# read in data, looks like sql dump, blah 
library(data.table) 
states <- fread(input = "./us_cities_and_states/states.sql", skip = 23, sep = "'", sep2 = ",", header = F, select = c(2, 4))

#read in data for cities
cities <- fread(input = "./us_cities_and_states/cities_extended.sql", header = F, sep = "'", sep2 = ",")

#matching the cities to only the states listed in "states"
a <- states[["V4"]]
b <- cities[["V4"]]
city_state <- match(b, a) %>%
  na.omit(city_state)

#finding the number of cities per state represented
city_count <- c(1:length(a))
for (i in 1:length(a)){
  city_count[i] <- length(which(city_state == i))
}

#making a table with the city count per state
city_count_table <- cbind(states, city_count)
print(city_count_table)

#this counts the number of a certain letter in a given word
letter_count <- data.frame(matrix(NA, nrow = length(a), ncol = 26))
get_count <- function(letter, state_name){
  d <- strsplit(state_name, "")[[1]]
  e <- letter
  f <- length(na.omit(match(d, e)))
  return(f)
}
```

