---
title: "STAT 5014 Homework 2"
author: "Samantha Sunshine"
date: "9/11/17"
output: pdf_document
---

# Problem 4

Version control can be beneficial in the classroom if I want to allow other people to work on a piece of code that I share with them. It would also help if I have made a mistake and realize I need to go back to a previous version of the code. Another benefit would be that I can access my files from multiple computers if I need to work on something and do not have my own computer.  

# Problem 5

## a. Sensory Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
url<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
Sensory_raw<-read.table(url, header=F, skip=1, fill=T, stringsAsFactors = F)
Sensory_tidy<-Sensory_raw[-1,]
Sensory_tidy_a<-filter(.data = Sensory_tidy,V1 %in% 1:10) %>%
  rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
Sensory_tidy_b<-filter(.data = Sensory_tidy,!(V1 %in% 1:10)) %>%
  mutate(Item=rep(as.character(1:10),each=2)) %>%
  mutate(V1=as.numeric(V1)) %>%
  select(c(Item,V1:V5))
Sensory_tidy<-bind_rows(Sensory_tidy_a,Sensory_tidy_b)
colnames(Sensory_tidy)<-c("Item",paste("Person",1:5,sep="_"))
Sensory_tidy<-Sensory_tidy %>%  
  gather(Person,value,Person_1:Person_5) %>%  
  mutate(Person = gsub("Person_","",Person)) %>%
  arrange(Item)
Sensory_table<-Sensory_tidy[1:20, ]
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(Sensory_table, caption="Sensory Data")
```

The first step was to create a table names Sensosry_raw that included the data from the url with no header and skipped the first line of original data.  
```{r, echo=TRUE}
Sensory_raw<-read.table(url, header=F, skip=1, fill=T, stringsAsFactors = F)
```

Then I created Sensory_tidy, which was the same as Sensory_raw but without the first row.  
```{r, echo=TRUE}
Sensory_tidy<-Sensory_raw[-1,]
```

The next line of code named a new dataset, Sensory_tidy_a, with the data from Sensory_tidy. This filtered the column V1 with all the numbers between 1 and 10, and it renamed the columns.  
```{r, echo=TRUE}
Sensory_tidy_a<-filter(.data = Sensory_tidy,V1 %in% 1:10) %>%
  rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
```

The next chunk of code created another dataset, Sensory_tidy_b, which took the data from column V1 that was not between 1 and 10. Then the mutate function made the "Item" column repeat the numbers 1 through 10, each two times. Column V1 is now numeric factors, and I selected the data in "Item", and columns V1 through V5.  
```{r, echo=TRUE}
Sensory_tidy_b<-filter(.data = Sensory_tidy,!(V1 %in% 1:10)) %>%
                    mutate(Item=rep(as.character(1:10),each=2)) %>%
                    mutate(V1=as.numeric(V1)) %>%
                    select(c(Item,V1:V5))
```

The following line combined the two datasets as rows, one on top of the other.
```{r,echo=TRUE}
Sensory_tidy<-bind_rows(Sensory_tidy_a,Sensory_tidy_b)
```

By renaming the column names of Sensory_tidy, I now have columns "Item", "Person_1", "Person_2", etc.  
```{r, echo=TRUE}
colnames(Sensory_tidy)<-c("Item",paste("Person",1:5,sep="_"))
```

The last chunk of code changed Sensory_tidy so that it created a column named "Person", which contained "Person_1", "Person_2", ... corresponding to each data point. The corresponding data points were gathered into the new column "value". The mutate function changed the "Person" column so that we substituted a blank space for "Person_", which left just the numbers, "1, 2, ...". The last step was to arrange the data by the "Item" column in ascending order.
```{r, echo=TRUE}
Sensory_tidy<-Sensory_tidy %>%  
        gather(Person,value,Person_1:Person_5) %>%  
        mutate(Person = gsub("Person_","",Person)) %>%
        arrange(Item)
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(summary(Sensory_tidy), caption="Sensory data summary")
```

## b. Long Jump Data

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
url<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
LongJump_raw<-read.table(url,header = F,skip = 1,fill = T,stringsAsFactors = F)
colnames(LongJump_raw)<-rep(c("V1","V2"),4)
LongJump_tidy<-rbind(LongJump_raw[ ,1:2],LongJump_raw[ ,3:4],LongJump_raw[ ,5:6],LongJump_raw[ ,7:8])
LongJump_tidy<-LongJump_tidy %>%
  filter(!(is.na(V1))) %>%
  mutate(YearCode=V1,Year=V1+1900,dist=V2) %>%
  select(-V1,-V2)
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(LongJump_tidy, caption="Long Jump Data")
```

The first step was to create a table named LongJump_raw from the url data with no header and by skipping the first line.
```{r, echo=TRUE}
LongJump_raw<-read.table(url,header = F,skip = 1,fill = T,stringsAsFactors = F)
```

Next, I renamed the column names of LongJump_raw as V1 and V2, repeated 4 times.
```{r, echo=TRUE}
colnames(LongJump_raw)<-rep(c("V1","V2"),4)
```

The next line of code named a new dataset, LongJump_tidy, which combined the data in LongJump_raw. Columns 1 and 2 were combined, as were 3 and 4, 5 and 6, and finally 7 and 8.
```{r, echo=TRUE}
LongJump_tidy<-rbind(LongJump_raw[ ,1:2],LongJump_raw[ ,3:4],LongJump_raw[ ,5:6],LongJump_raw[ ,7:8])
```

The last chunk of code changed LongJump_tidy to filter the data in column V1 that is not missing. The mutate function created new columns YearCode, which was the same as V1, Year, which added 1900 to V1 so it was an actual year, and dist, which was the same as V2. Then I selected the data set without columns V1 and V2.
```{r, echo=TRUE}
LongJump_tidy<-LongJump_tidy %>%
  filter(!(is.na(V1))) %>%
  mutate(YearCode=V1,Year=V1+1900,dist=V2) %>%
  select(-V1,-V2)
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(summary(LongJump_tidy), caption="Long Jump data summary")
```

## c. Brain and Body Data
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
url<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
BrainBody_raw<-read.table(url,header=F,skip = 1,fill = T,stringsAsFactors = F)
colnames(BrainBody_raw)<-rep(c("Brain","Body"),3)
BrainBody_tidy<-rbind(BrainBody_raw[ ,1:2],BrainBody_raw[ ,3:4],BrainBody_raw[ ,5:6])
BrainBody_tidy<-BrainBody_tidy %>%
  filter(!(is.na(Brain)))
BrainBody_table<-BrainBody_tidy[1:20, ]
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(BrainBody_table, caption="Brain and Body Data")
```

First, I created a table named BrainBody_raw from the url data, which had no header, and skipped the first line of data.
```{r,echo=TRUE}
BrainBody_raw<-read.table(url,header=F,skip = 1,fill = T,stringsAsFactors = F)
```

The next step was to rename the columns as "Brain" and "Body" repeated 3 times.
```{r,echo=TRUE}
colnames(BrainBody_raw)<-rep(c("Brain","Body"),3)
```

I created a new dataset named BrainBody_tidy, which combined the columns from BrainBody_raw. Columns 1 and 2 were now combined, as were columns 3 and 4, and columns 5 and 6.
```{r,echo=TRUE}
BrainBody_tidy<-rbind(BrainBody_raw[ ,1:2],BrainBody_raw[ ,3:4],BrainBody_raw[ ,5:6])
```

Finally, BrainBody_tidy was filtered by the data that was not missing from the "Brain" column.
```{r,echo=TRUE}
BrainBody_tidy<-BrainBody_tidy %>%
  filter(!(is.na(Brain)))
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(summary(BrainBody_tidy), caption="Brain/Body weight data summary")
```

## d. Tomato Data
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
url<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
Tomato_raw<-read.table(url,header = F,skip = 2,fill = T,stringsAsFactors = F,comment.char = "")
Tomato_tidy<-Tomato_raw %>%
  separate(V2,into=paste("C10000",1:3,sep="_"),sep=",",remove=T,extra="merge") %>%
  separate(V3,into=paste("C20000",1:3,sep="_"),sep=",",remove=T,extra="merge") %>%
  separate(V4,into=paste("C30000",1:3,sep="_"),sep=",",remove=T,extra="merge") %>%
  mutate(C10000_3=gsub(",","",C10000_3)) %>%
  gather(Clone,value,C10000_1:C30000_3) %>%
  mutate(Variety=V1,Clone=gsub("C","",Clone)) %>%
  mutate(Variety=gsub("\\\\#"," ",Variety)) %>%
  separate(Clone,into=c("Clone","Replicate")) %>%
  select(-V1,Variety,Clone,value) %>%
  arrange(Variety)
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(Tomato_tidy, caption="Tomato Data")
```

First, I made a table named Tomato_raw from the url data, which had no header and skipped the first two lines.
```{r,echo=TRUE}
Tomato_raw<-read.table(url,header = F,skip = 2,fill = T,stringsAsFactors = F,comment.char = "")
```

Tomato_tidy is the new dataset that represents Tomato_raw, except I separated V2 into 3 columns, "C10000_1", "C10000_2", and "C10000_3". I also removed the comma that was originally separating the data.  
I did the same thing for V3 and V4, which split into 3 columns each, "C20000_1", "C20000_2", and "C20000_3", and "C30000_1", "C30000_2", and "C30000_3", respectively.  
The mutate function eliminated the extra comma in C10000_3.  
The gather function created a new column named "Clone", which contained "C10000_1", "C10000_2", ... corresponding to each data point. The corresponding data points were arranged into a new column named "value".  
Mutate created a new column, "Variety", which was the same as column V1. It also substituted a blank space for "C" in the "Clone" column, which left "10000_1", "10000_2", etc. 
Mutate also substituted a blank space for "\\#" in the "Variety" column.  
The separate function broke the "Clone" column into two by splitting "10000_1" into "10000" in the "Clone" column and "1" in the new "Replicate" column.  
The last two parts were to select the data without column V1, and with columns Variety, Clone, and value. Then I arranged the data by Variety.  
```{r,echo=TRUE}
Tomato_tidy<-Tomato_raw %>%
  separate(V2,into=paste("C10000",1:3,sep="_"),sep=",",remove=T,extra="merge") %>%
  separate(V3,into=paste("C20000",1:3,sep="_"),sep=",",remove=T,extra="merge") %>%
  separate(V4,into=paste("C30000",1:3,sep="_"),sep=",",remove=T,extra="merge") %>%
  mutate(C10000_3=gsub(",","",C10000_3)) %>%
  gather(Clone,value,C10000_1:C30000_3) %>%
  mutate(Variety=V1,Clone=gsub("C","",Clone)) %>%
  mutate(Variety=gsub("\\\\#"," ",Variety)) %>%
  separate(Clone,into=c("Clone","Replicate")) %>%
  select(-V1,Variety,Clone,value) %>%
  arrange(Variety)
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(summary(Tomato_tidy), caption="Tomato data summary")
```

# Problem 6

```{r, echo=TRUE, include=FALSE}
library(swirl)
.datapath <- file.path(path.package('swirl'), 'Courses', 'R_Programming_E', 'Looking_at_Data', 'plant-data.txt') 
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")
.cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol')
plants <- plants[, !(names(plants) %in% .cols2rm)]
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period', 'Foliage_Color', 'pH_Min', 'pH_Max', 'Precip_Min', 'Precip_Max', 'Shade_Tolerance', 'Temp_Min_F')
plants_tidy<-plants %>%
  mutate(pH_Total=(pH_Min+pH_Max)) %>%
  mutate(pH_Average=pH_Total/2) %>%
  select(Scientific_Name,Foliage_Color,pH_Average) %>%
  na.omit %>%
  arrange(pH_Average)
plants_table<-plants_tidy[1:20,]
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(plants_table, caption="Plants Data")
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(summary(plants_tidy), caption="Plants Data Summary")
```

```{r, echo=FALSE,include=TRUE,comment=NA}
lm(plants_tidy)
```

# Problem 7

```{r,echo=FALSE, include=TRUE,comment=NA}
Car_Gebreken_raw<-read.csv("C:/Users/ssuns/Documents/2017 Fall/Statistical Programming/STAT_5014_Homework/Open_Data_RDW__Gebreken.csv",stringsAsFactors = F, nrows=200, header=T,quote = '"')
Car_Geconstat_raw<-read.csv("C:/Users/ssuns/Documents/2017 Fall/Statistical Programming/STAT_5014_Homework/Open_Data_RDW__Geconstateerde_Gebreken.csv", stringsAsFactors = F, nrows=200, header=T)
Car_Person_raw<-read.csv("C:/Users/ssuns/Documents/2017 Fall/Statistical Programming/STAT_5014_Homework/Personenauto_basisdata.csv",stringsAsFactors = F, nrows=200, header=T)
Car_Gebreken_raw.colclass<-sapply(Car_Gebreken_raw,class)
Car_Geconstat_raw.colclass<-sapply(Car_Geconstat_raw,class)
Car_Person_raw.colclass<-sapply(Car_Person_raw,class)
print("Gebreken")
print(Car_Gebreken_raw.colclass)
print("Geconstat")
print(Car_Geconstat_raw.colclass)
print("Personen")
print(Car_Person_raw.colclass)
Car_Gebreken_table<-Car_Gebreken_raw[1:20, ]
Car_Geconstat_table<-Car_Geconstat_raw[1:20, ]
Car_Person_table<-Car_Person_raw[1:20, ]
```

```{r, echo=FALSE,include=TRUE}
knitr::kable(Car_Gebreken_table, caption="Gebreken Data")
knitr::kable(summary(Car_Gebreken_table), caption="Gebreken Summary Data")
knitr::kable(Car_Geconstat_table, caption="Geconstat Data")
knitr::kable(summary(Car_Geconstat_table), caption="Geconstat Summary Data")
knitr::kable(Car_Person_table, caption="Person Data")
knitr::kable(summary(Car_Person_table), caption="Person Summary Data")
```

